;[---
title: "covid 19"
author: "zaira gomez"
date: "10/19/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```
The results are 1831 females are live and 1411 men aalive more women than  men are alive 
## We create two bar charts - one based on frequency and the other on percent

## Boxplot 

```{r}
barchartGC(~result + Sex,data=kRdata, main ="Frequency Bar Chart")
```




The bar chart is displaying in total frequency terms our data.Clearly there are a lot more dead men than women as is reflected by the size of the graph .

```{r}
barchartGC(~result + Sex,data=kRdata, type = "percent",main ="Frequency Bar Chart")
```
69% of women 41 % are men still alive in the dead groupo 40% of womenare dead and 62% of men are dead .


## Numerical Results 
```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

Of the females who are alive 56% are females and 44% are male .Women who are dead is 44% of females and dead males are 64% .Males are dying a lot more often then females .

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

More men are dead thn live compared to women who have less deaths .




## Conclusions
We concluded that men are most likely to die the ratio is 2.3 .The males have a higher percentage of deaths than the females .


