---
title: "covid 19"
author: "Dr K"
date: "5/2/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(tigerstats)
require(lubridate)
kdata <- read.csv(file = 'PatientInfo.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex


```{r, include=FALSE}
kRdata <- filter(kdata,sex !="") %>% mutate(Sex=if_else(sex=="female","F","M")) %>% mutate(result=(if_else(state=="deceased","dead","alive" ))) 
kRdata$result <- as.factor(kRdata$result)
kRdata$Sex <- as.factor(kRdata$Sex)
kRdata <- as.tibble(kRdata)
```

Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, state))

```

## We create two bar charts - one based on frequency and the other on percent

```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata,type="percent")
```

## We see from our row percent numerical data a disparity between men and women  

```{r}

table1 <- xtabs(~result + Sex,data=kRdata)
rowPerc(table1)
colPerc(table1)

```


## Inferential results are significant above the .005 level

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
## Explanation of Fisher exact test

The odds ratio is the odds of a Female being alive compared to the odds of a male being alive 
in this case females are 2.3 times more likely to survive than are males.   Alternatively it is the odds of a male death divided by the odds of a female death.  Basically bad news for men.

# Question 2 what is the age distribution of those who contracted Covid contrasted with those who died

Since we have the birth year, we wii simply plot the difference between that and 2020
we will form two plots one for everyone ane one for just those who died

```{r}
age = 2020-na.omit(kRdata$birth_year)
plot(density(age))
kDead <- filter(kRdata,result=="dead")
ageDead =2020-na.omit(kDead$birth_year)
plot(density(ageDead))

```

# Question 3  When did the people get covid


```{r}
# create date with lubridate
kRdata$confirmed <- ymd(kRdata$confirmed_date)
ggplot(kRdata,aes(confirmed)) +
geom_freqpoly(binwidth=2,na.rm=TRUE) +
ggtitle("Cases confirmed by Date")
```

# Question 4 Time it goes on

```{r}
kRdata$kRdataRelease <- ymd(kRdata$released_date)
kRdata$kRdataDead <- ymd(kRdata$deceased_date)
```


```{r}
goesOnR <-  as.numeric(kRdata$kRdataRelease) - as.numeric(kRdata$confirmed) 
plot(density(na.omit(goesOnR)))
gor <- na.omit(goesOnR)
hist(gor,main=paste("Time to Recover"))
goesOnD <-  as.numeric(kRdata$kRdataDead) - as.numeric(kRdata$confirmed) 
plot(density(na.omit(goesOnD)))
goD <-na.omit(goesOnD)
hist(goD,main=paste("Time till Death"))
```

  
#### see https://data.library.virginia.edu/working-with-dates-and-time-in-r-using-the-lubridate-package/
